---
title: Graham Norton's Guests
author: Connor Kirkpatrick
date: '2017-10-16'
slug: graham-norton-s-guests
categories: []
tags: []

---

# The Guests of the Graham Norton Show

Sick, lying on the sofa trying to pass an evening. I found myself in desperate need of something easy to watch as I waited for the headache, sore throat and fever to leave me. Of the shows  I did want to watch, all are shows I watch with my girlfriend. Even sick, I don't think I could get away with that level of betrayl!

So I put of Graham Norton. As my girlfriend describes it, "It's okay if you miss a bit". Perfect for when you're half dozing.

Nicole Kidman was the first guest on this weeks show. At the start, she says "I think I've been on here the most of anyone". That was like drugs to a sniffer dog, even to my tired mind. Now that I'm feeling a little better, let's see if we can answer it.

[Wikipedia has a page detailing every episode and it's guests](https://en.wikipedia.org/wiki/List_of_The_Graham_Norton_Show_episodes). We can extract the information from this to get our data for analysis.



```{r warning=FALSE, message=FALSE}
library(rvest)
library(purrr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)

page <- read_html("https://en.wikipedia.org/wiki/List_of_The_Graham_Norton_Show_episodes")

#list all of the selectors
selector_guests <- ".summary"
selector_episode_no <- ".vevent td:nth-child(2)"
selector_overall_episode_no <- ".vevent th"
selector_air_date <- ".vevent td:nth-child(4)"

selectors <- c(air_date = selector_air_date, 
               guests = selector_guests, 
               episode_no = selector_episode_no, 
               overall_episode_no = selector_overall_episode_no)

# Create general function to extract text from page
extract_graham_norton <- function(site, selector) {
  site %>% html_nodes(selector) %>% html_text()
}

lists_mania <- map_df(selectors, ~extract_graham_norton(page, .x))

df <- bind_cols(lists_mania)

head(df)
```


We have the data in it's rough form, but we'll need to extract the names into something we can work with. Currently they are in the form of `"guest1, guest2, ..., guest(n-1) and guest(n)"`. We'll need to split the string on  `","` and `"and"`, then unnest our new dataframe to create a record for every guest appearance on the show.

```{r warning=F, message=F}
tidy_df <- df %>%
  
  #Remove "compilation show episodes. These have no guests"
  filter(guests != "Compilation show") %>%
  
  # Extract the air date and parse it. Wikipedia has a strange format for these dates
  mutate(air_date_tidy = str_extract(air_date, "(.{1,2}\\W[A-z]+\\W\\d{4})") %>% dmy(),
         
         # Split the listing the guests on "," or "and"
         guests = str_split(guests, ",|\\band\\b")) %>%
  
  tidyr::unnest() %>%
  
  # Tidy up the names so that they are in a uniform format
  mutate(guests = str_to_lower(guests) %>% str_replace_all("pilot|\\(|\\)", "") %>% str_trim())

#filename <- paste("output-data/", "graham-norton-guests_", Sys.Date(), ".csv", sep = "")
#write.csv(tidy_df, file = filename)

head(tidy_df)
```

Our data is now fit for purpose. Is Nicole Kidman the most common guest on the show? Let's check the top 5


```{r}

tidy_df %>%
  count(guests) %>%
  arrange(desc(n)) %>%
  top_n(5, n)


```

She's no where near! How many times has she been on the show?

```{r}
tidy_df %>%
  filter(guests =="nicole kidman") %>%
  count()


```



## Bonus


### Histogram of Guest frequency
```{r}
tidy_df %>%
  count(guests) %>%
  ggplot(aes(x = n)) + geom_bar() + labs(x = "Occurence", y = "Frequency", title = "A barchart to show the number of times each guest has appeared on the show ", subtitle = "Most guests only appear once")


tidy_df %>%
  count(guests) %>%
  group_by(n) %>%
  summarise(count = n())
  

```


### Number of distinct guests

How many guests has Graham Norton had?

Graham has had `{r} nrow(tidy_df)` guest appearances. How many of these were distinct guests have made an appearance?

```{r}
# Total Guest appearances
nrow(tidy_df)

# Distinct number of guests
tidy_df %>%
  distinct(guests) %>%
  count()
```

### Number of Guests Per Epidsode

```{r}
tidy_df %>%
  count(overall_episode_no) %>%
  arrange(desc(n)) %>%
  top_n(10, wt = n)
```

What's so special about epidsode 206? It appears to be a New Years Eve special 2013
```{r}

tidy_df %>%
  filter(overall_episode_no == 206) 
```


```{r}
tidy_df %>%
  group_by(overall_episode_no) %>%
  count() %>% 
  ungroup() %>%
  summarise(mean_no_of_guests = mean(n),
            median_no_of_guests = median(n),
            max_no_of_guests = max(n),
            min_no_of_guests = min(n))

```
