---
title: How does `+` in ggplot work?
author: Connor Kirkpatrick
date: '2017-09-06'
slug: how-does-in-ggplot-work
categories: []
tags:
  - ggplot
---

I am a big fan of `ggplot2`. As one of the most popular R packages, I'm clearly not alone. As well as making great plots, it's given me a deeper concept of what a plot is, and how data maps to a geom (in gg terms).

One of the most recognisable parts of the package is the concept of adding layers to plots. We'll use the `steak_survey` dataset from the `fivethirtyeight` package. (Check out their website for some interesting data related stories!)



```{r firstplot}
library(ggplot2)
library(fivethirtyeight)

steak <- fivethirtyeight::steak_survey

# Create example plot

ggplot(steak, aes(region, steak_prep, col = steak_prep)) + geom_jitter() + coord_flip()
```


The code to create this plot literally adds layers, using `+`. But how does the `+` work? It's not a number.., so what magic is going on? Can we remove layers with `-`?


To get our answer, there are two parts to understand. 

## Part One - What exactly is `+`?

It seems pretty obvious. Everyone will recognise the `+` symbol from an early age. But what is it exactly in R?

```{r}

1 + 2
# 3
```

`+` is in fact, just another function. The ability to perform `x + y` is syntactic sugar for the language. It's much easier than typing it's functional form as shown in the example below.

```{r}
class(`+`)

1 + 2

`+`(1, 2)

```

So when we call `ggplot() + geom_*`, we're actually just calling another function like any other. We could recreate our above graph with the code below, as it's equivalent (apart from the missing `coord_flip`).

```{r secondplot}
`+`(ggplot(steak, aes(region, steak_prep, col = steak_prep)), geom_jitter())

```

So that explains what `+` is, but we still don't know how `R` knows what the sum of a `ggplot() + geom_*` call is.

## Part Two - Generic Functions

To get out answer, we'll take a small detour along the way.

Let's look at `print()`

```{r}
print("abcdef") # works as expected
print(123) # Still works as expected

```

Not very interesting so far, so let's break it.

```{r}

time <- Sys.time() # return current time
typeof(time) # double? But it's a date-time!?

print(time) # This prints the date though..
```

As you can see, the time, is not strictly a time. It's actually a double, counting the number of seconds since the Unix epoch. But when we print it, it returns a pretty date? Very odd.

```{r}
print(as.numeric(time)) # That's what it actually is
```
So how does it know to print something else? The answer- Generic functions. When you call print, a bit more goes on behind the scenes that you might initially realise. `print` is known as a generic method. Depending on the class of the object is it trying to print, a different print function will be called! You can see all of the methods for print using the `methods()` function.

```{r}
typeof(print)
class(print)
head(methods(print))
length(methods(print))

```

When you print the time, it checks to see the class of the object. Our time object has the classes of `POSIXct` and `POSIXt`. It then calls the `print.POSIXct()` method, to nicely format out time.

### But what does this have to do with `+`?

You may have already made the leap, but as we discussed earlier, `+` is just another function. As such, we could assign it another method. In this case, it's as below.

```{r}

methods(`+`)

# Note +.gg* - This is the method we are after!
```

So in summary, ggplot2 clever identifies that `+` is a function, and then assigns it a new method, which is then uses to create beautiful plots! 

For a bit more information, check out the following

1. The help page for the ggplot `+` method.

```{r}
help(`+.gg`)
```

1. For more about generic functions, check out [Advanced R, by Hadley Wickham](http://adv-r.had.co.nz/),  who also authored the `ggplot` package!
